\texttt{ensmallen}: a flexible C++ library for efficient function optimization

### 2018-10-17 doc/mlsys_paper/paper.tex
sharper and shorter conclusion

Lastly, we demonstrate the easy pluggability in {\tt ensmallen}
for using various optimizers on the same task.
Using a version of \texttt{\small LinearRegressionFunction} from Sec.~\ref{sec:linreg_example}
adapted for separable differentiable optimizers,
we run six optimizers with default parameters in just 7 lines of code,
as shown in Fig.~\ref{fig:learning_curve}(a).
Applying these optimizers to the {\tt YearPredictionMSD}
dataset from the UCI repository~\cite{ucimlrepository}
yields the learning curves shown in Fig.~\ref{fig:learning_curve}(b).
Any other optimizers for separable differentiable objective
functions can be dropped into place in the same manner.

@@

ens::StandardSGD<>().Optimize(lrf, sgdModel);
ens::Adam().Optimize(lrf, adamModel);
ens::AdaGrad().Optimize(lrf, adagradModel);
ens::SMORMS3().Optimize(lrf, smorms3Model);
ens::SPALeRASGD().Optimize(lrf, spaleraModel);
ens::RMSProp().Optimize(lrf, rmspropModel);
\end{minted}
\end{minipage}
\end{adjustbox}
\hfill
\begin{minipage}{0.49\textwidth}
  \centering
  \includegraphics[width=\textwidth,height=0.45\textwidth]{experiments/learning_curves_crop.png}
\end{minipage}
\end{minipage}
\begin{minipage}[t]{1\textwidth}
\begin{minipage}{0.50\textwidth}
\centering
(a)
\end{minipage}
\hfill
\begin{minipage}{0.49\textwidth}
\centering
(b)
\end{minipage}
\end{minipage}
\caption
  {
  (a): example code for using 
  six {\tt ensmallen} optimizers with default parameters
  to optimize a linear regression function;
  (b): corresponding learning curves
  on the {\tt YearPredictionMSD} dataset~\cite{ucimlrepository}
  over 5 epochs of training.
  The optimizers can be tuned for better performance.
  }
\label{fig:learning_curve}
\end{figure}

@@

We have described {\tt ensmallen}, a flexible C++ library for function
optimization that provides an easy interface for users to implement and optimize
their desired objective functions.  Many types of functions can be optimized,
including separable and constrained functions.  The library is fast due to its
use of template metaprogramming and the Armadillo matrix library for linear
algebra.  In addition, template metaprogramming is used to make user
implementation easier by automatically generating missing methods.

The library provides many pre-built optimizers (including numerous variants
of SGD and Quasi-Newton optimizers).
We aim to expand the library with further optimization techniques
as the need arises.  Since {\tt ensmallen} is open source,
external contributions to the codebase are welcome.

% GPU support for optimization is available either through the use of the NVBLAS
% library~\cite{nvidia2015} or the forthcoming Bandicoot C++ GPU matrix library.
%% Conrad: is it ok if we mention that here?  It's a really nice selling point
%% and I am sure at least one reviewer will be thinking about GPUs.

%% CS: we don't have much room, so I moved the mention of NVBLAS to the intro.
%% CS: i think it flows better there.
%% CS: mentioning it here in the conclusion seems a bit left-field.
%% CS: we also don't have a good reference for Bandicoot yet

{\tt ensmallen} is already in use as the optimization toolkit of the {\tt
mlpack} machine learning library~\cite{mlpack2018}.  For more information on the
toolkit, see the website and documentation at \url{https://www.ensmallen.org/}.

>>>>

Lastly, we demonstrate the easy pluggability in {\tt ensmallen}
for using various optimizers on the same task.
Using a version of \texttt{\small LinearRegressionFunction} from Sec.~\ref{sec:linreg_example}
adapted for separable differentiable optimizers,
we run six optimizers with default parameters in just 7 lines of code,
as shown in Fig.~\ref{fig:learning_curve}(a).
Applying these optimizers to the \texttt{\small YearPredictionMSD}
dataset from the UCI repository~\cite{ucimlrepository}
yields the learning curves shown in Fig.~\ref{fig:learning_curve}(b).
Any other optimizers for separable differentiable objective
functions can be dropped into place in the same manner.

@@

ens::StandardSGD<>().Optimize(lrf, sgdModel);
ens::Adam().Optimize(lrf, adamModel);
ens::AdaGrad().Optimize(lrf, adagradModel);
ens::SMORMS3().Optimize(lrf, smorms3Model);
ens::SPALeRASGD().Optimize(lrf, spaleraModel);
ens::RMSProp().Optimize(lrf, rmspropModel);
\end{minted}
\end{minipage}
\end{adjustbox}
\hfill
\begin{minipage}{0.49\textwidth}
  \centering
  \includegraphics[width=\textwidth,height=0.45\textwidth]{experiments/learning_curves_crop.png}
\end{minipage}
\end{minipage}
\begin{minipage}[t]{1\textwidth}
\begin{minipage}{0.50\textwidth}
\centering
(a)
\end{minipage}
\hfill
\begin{minipage}{0.49\textwidth}
\centering
(b)
\end{minipage}
\end{minipage}
\caption
  {
  \footnotesize
  (a): example code for using 
  six {\tt ensmallen} optimizers with default parameters
  to optimize a linear regression function;
  (b): corresponding learning curves
  on the \texttt{\small YearPredictionMSD} dataset~\cite{ucimlrepository}
  over 5 epochs of training.
  The optimizers can be tuned for better performance.
  }
\label{fig:learning_curve}
\end{figure}

@@

We overviewed {\tt ensmallen}, a flexible C++ library for function
optimization that provides an easy interface for the implementation and optimization
of user-defined objective functions.  Many types of functions can be optimized,
including separable and constrained functions.
The library provides many pre-built optimizers (including numerous variants
of SGD and Quasi-Newton optimizers).
The library internally exploits template metaprogramming
to maximise opportunities for efficiency gains,
as well as 
to make the implementation of objective functions easier
by automatically generating missing methods.

% {\tt ensmallen} is already in use as the optimization toolkit of the {\tt
% mlpack} machine learning library~\cite{mlpack2018}.

We aim to expand the library with further optimization techniques
as the need arises.  Since {\tt ensmallen} is open source,
external contributions to the codebase are welcome.
For more information on the
toolkit, see the website and documentation at \url{https://www.ensmallen.org/}.
